---
title: "A Report on Lake Ontario's Microbes"
author: "Autumn Pereira"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#this is global setup,so we can set echo =F if we
#want it to not output the code in the knit document
knitr::opts_chunk$set(echo = TRUE)
```

#### Prepare our environment!

```{r packages}
#load libraryes
library(tidyverse)
```

```{r load-data}
#load in lake ontario microbial community data
sample_and_taxon <- read_csv("data/sample_and_taxon.csv")

```

#### Lake Ontario Microbial Abundance vs Temperature

```{r cells-vs-temp-plot}
ggplot(data=sample_and_taxon)+
  geom_smooth(aes(x=temperature, y=cells_per_ml/1000000), method= "lm", color = "grey30")+
  geom_point(aes(x= temperature, y= cells_per_ml/1000000, color = env_group), size = 0.75)+
  labs(x= "Temperature (°C)", y = "Cell Abundance (million/mL)", title = "Relationship Between Cell Abundances and Temperature in Lake Ontario")+
  theme_bw()+
  theme(legend.position= "bottom", legend.title=element_blank())
```

It seems that there is a significant positive linear relationship between cell abundances and temperature in Lake Ontario, and that deep samples are the coldest and have the fewest cells

```{r data-summary, include = FALSE}
#how many samples do i have?
n_samples <- sample_and_taxon %>%
  #only pick out sample ID column
  select(sample_id) %>%
  #number of distinct names in the column
  n_distinct()

#calculate min and max temps
#min
min_temp <- sample_and_taxon %>%
  summarize(round(min(temperature), digits=1)) %>%
  pull

#max
max_temp <- sample_and_taxon %>%
  summarize(round(max(temperature), digits=1))%>%
  pull
```

The total number of samples is `r n_samples`. For this set of samples, temperature ranged from a minimum of `r min_temp`°C to `r max_temp`°C.
